Implicit assumptions and their impact on photometric redshift PDF performance in the context of LSST

### 2018-11-21 paper/main.tex
change title (temp) and few edits

\title{An assessment of photometric redshift PDF performance in the context of LSST}
% \title{Photometric redshift uncertainties for LSST \romannumeral{1}: A unified methodology for assessing photo-$z$ PDF performance}

@@

The authors would like to thank their LSST-DESC publication review committee for comments that improved the paper draft.

@@

In addition to packages cited in the text, analyses performed in this paper used the following software packages: \textsc{Numpy} and \textsc{Scipy} \citep{numpyscipy}, \textsc{Matplotlib} \citep{matplotlib}, \textsc{Seaborn} \citep{seaborn},

>>>>

\title{Implicit assumptions and their impact on photometric redshift PDF performance in the context of LSST}
% \title{Photometric redshift uncertainties for LSST \romannumeral{1}: A unified methodology for assessing photo-$z$ PDF performance}

@@

The authors would like to thank their LSST-DESC publication review committee.

@@

### 2018-11-21 paper/tex/introduction.tex
change title (temp) and few edits

The unprecedented sample size of LSST galaxies, expected to number several billion for the main cosmological sample, necessitates stringent constraints on photo-$z$ accuracy if systematic errors are not to dominate the statistical errors.
The LSST Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} lists the photometric redshift goals for a magnitude limited sample with $i<25$ as: root-mean-square error with a goal of $\sigma_z<0.02(1+z)$; $3\sigma$ ``catastrophic outlier'' rate below $10\%$; bias below $0.003$\footnote{
Note that at the time the SRD was written, these goals were stated in terms of a photo-$z$ point estimate for each galaxy, as was standard in many previous studies, while in this paper we emphasize the importance of using a full photo-$z$ PDF.}.  The LSST Dark Energy Science Collaboration (LSST-DESC) developed a separate Science Requirements Document \citep{Mandelbaum:2018}, which forecasts the constraining power of five cosmological probes using somewhat conservative assumptions to define requirements on systematic errors for several measurements.  These include even more stringent requirements on photometric redshift performance than those included in the LSST SRD, though the initial LSST-DESC requirements are defined for tomographic bin populations.  The tremendous size of LSST's galaxy catalogue will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
The inclusion of these populations introduce major physical degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in the populations covered in shallower large area surveys like the Sloan Digital Sky Survey \citep[SDSS,][]{York:00} and the Two Micron All Sky Survey \citep[2MASS,][]{Skrutskie:06}.  These issues are not unique to LSST, and are present in Stage III Dark Energy surveys; however,
in order to meet the demanding error budgets of Stage IV projects such as LSST and LSST-DESC it will be necessary to fully characterize those degeneracies wherein multiple redshift solutions have comparable likelihood to per cent level accuracy.

@@

For cosmological measurements, certain science cases require redshift information on individual objects, e.~g.~identification of host galaxy redshift for supernova classification, or identifying potential cluster membership.
Other science cases seem to need only ensemble redshift information; for instance many current cosmic shear techniques require only the overall redshift distribution $N(z)$ for tomographic redshift samples.  However, recent data-driven techniques employing hierarchical Bayesian or Gaussian Process methods have emerged that calibrate redshift distributions using individual $p(z)$ estimates \citep[e.~g.~][]{Sanchez:2018}.  These methods assume that the $p(z)$ for each galaxy is an accurate PDF, and such methods break down if this assumption is invalid.  Thus, even methods that seem to need only ensemble $N(z)$ may actually require accurate $p(z)$ in order to meet stringent survey requirements.
%In the case of the multiple types of probes of cosmology enabled by the
%LSST cosmology sample of several billion galaxies, the number of redshift bins and their photo-$z$ requirements vary with the specific probe;  2-point angular correlations benefit from many bins, while weak lensing probes do not (due to the wide lensing kernel).
Large photometric surveys such as LSST must develop algorithms that simultaneously meet the needs of all science cases.
In order to meet these ambitious goals for photo-$z$ accuracy, every aspect of photo-$z$ estimation will have to be optimized: the algorithms employed, both template and machine-learning based (both in design and implementation); the spectroscopic data used as a training set for machine learning algorithms or to estimate template sets and train Bayesian priors; and probabilistic catalogue compression schemes that balance information retention against limited storage resources.

There are numerous techniques for deriving photo-$z$ PDFs from photometry, yet no one method has yet been established as clearly superior.
Quantitative comparisons of photo-$z$ methods have been made before.
The Photo-$z$ Accuracy And Testing \citep[PHAT,][]{Hildebrandt:10} effort focused on point estimates derived from many photometric bands.  \citet{Rau:2015} introduced a new method for improving redshift PDFs using an ordinal classification algorithm.
DES compared several codes for point estimates \citep{Sanchez:14} and a summary statistic of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}.  
This paper is distinguished by its focus on metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product.
The results presented in this work are
%part of the Data Challenge 1 (DC1)
a major focus of the Photometric Redshift working group of the LSST-DESC.
This work is laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}} as one of the critical activities to be completed in preparation for dark energy science analysis on the first year LSST data.
This is the first of multiple papers by the working group, which will grow in sophistication.
In this initial paper we focus on evaluating the performance of photometric redshift codes and PDF-based performance metrics in the presence of representative training sets.
Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.  We note that these initial tests are a necessary requirement for photo-$z$ codes that will be used in cosmological analyses; however, meeting these requirements is only the first stage in the process of accounting for incomplete training samples, object blending,  the robustness of estimated redshift distributions when combining redshift estimates in tomographic bins, and other complications that realistic samples will encounter in estimating photometric redshifts.  This can be thought of as an initial test under near perfect conditions before further complexities are added in future papers.

Before moving forward, we must address how the best methods may be unique to the performance metrics and science cases considered and what distinguishes photo-$z$ PDFs of different methods from one another.
Though photo-$z$ PDFs are often written simply as $p(z)$, the PDF itself must be an interim posterior distribution $p(z | d, I)$, the probability of redshift conditioned on photometric data $d$ that has actually been observed and the prior information $I$ that guides how a redshift is extracted from the photometry.
If we run multiple photo-$z$ codes on a single dataset, the photo-$z$ interim posteriors will not be identical because each code is based on assumptions in the form of an interim prior --- these assumptions form the premise for photo-$z$ estimation as a whole and are the only way to introduce differences in estimates of what would otherwise be a shared photo-$z$ posterior $p(z | d)$ regardless of the code used to obtain it.
Though explicit knowledge of the interim prior is necessary to use photo-$z$ interim posteriors self-consistently in physical inference, the interim prior of a particular methodology is often implicit and not necessarily shared among all galaxies in the catalogue.

This paper therefore aims (1) to constrain the impact of the interim prior $I$ by separating it into a component $I_{H}$ representing the method itself and a component $I_{D}$ representing physical information, such as a training set or SED template library and (2) to present a procedure for evaluating the performance of photo-$z$ codes in generic tests that may include many more systematics in the interim prior $I$.
In order to isolate the effects encapsulated by $I_{H}$ of variation between codes from issues with the training set or template library encapsulated by $I_{D}$, we use an identical set of simulated galaxies for every code and construct a template library and training sample that are {\it complete and representative} and shared among all codes; that is, our training sample for machine learning codes is drawn from the same underlying galaxy population as our test set, with no additional selections, and the SED library used for template-based codes is the same as the one used to generate the photometric data.
We explore a number of performance metrics in this paper, not to make a conclusion regarding the superiority or even relative favorability of each code but to establish a method for comparing photo-$z$ PDFs derived by different methods.
These test conditions set the stage for addressing in a future paper the crucial issue of incomplete and non-representative prior information.

>>>>

The unprecedented sample size of LSST galaxies, expected to number several billion for the main cosmological sample, necessitates stringent constraints on photo-$z$ accuracy if systematic errors are not to dominate the statistical errors.
The LSST Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} lists the individual galaxy photometric redshift goals for a magnitude limited sample with $i<25$ as: root-mean-square error with a goal of $\sigma_z<0.02(1+z)$; $3\sigma$ ``catastrophic outlier'' rate below $10\%$; bias below $0.003$\footnote{
Note that at the time the SRD was written, these goals were stated in terms of a photo-$z$ point estimate for each galaxy, as was standard in many previous studies, while in this paper we emphasize the importance of using a full photo-$z$ PDF.}.  The LSST Dark Energy Science Collaboration (LSST-DESC) developed a separate Science Requirements Document \citep{Mandelbaum:2018}, which forecasts the constraining power of five cosmological probes using somewhat conservative assumptions to define requirements on systematic errors for several measurements.  These include even more stringent requirements on photometric redshift performance than those included in the LSST SRD, though most of the initial LSST-DESC requirements are defined in terms of tomographic bin populations rather than on individual object redshifts.  The tremendous size of LSST's galaxy catalogue will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
The inclusion of these populations introduce major physical degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in the populations covered in shallower large area surveys like the Sloan Digital Sky Survey \citep[SDSS,][]{York:00} and the Two Micron All Sky Survey \citep[2MASS,][]{Skrutskie:06}.  These issues are not unique to LSST, and are present in Stage III Dark Energy surveys; however,
in order to meet the demanding error budgets of Stage IV projects such as LSST and LSST-DESC it will be necessary to fully characterize those degeneracies wherein multiple redshift solutions have comparable likelihood to per cent level accuracy.

@@

For cosmological measurements, certain science cases require redshift information on individual objects, e.~g.~identification of host galaxy redshift for supernova classification, or identifying potential cluster membership.
Other science cases seem to need only ensemble redshift information; for instance many current cosmic shear techniques require only the overall redshift distribution $N(z)$ for tomographic redshift samples.  However,  even such cases require individual object redshift estimates for portions of the analysis, for example in determining galaxy intrinsic alignments in weak lensing samples.  In addition, recent data-driven techniques employing hierarchical Bayesian or Gaussian Process methods have emerged that calibrate redshift distributions using individual $p(z)$ estimates \citep[e.~g.~][]{Sanchez:2018}.  These methods assume that the $p(z)$ for each galaxy is an accurate PDF, and such methods break down if this assumption is invalid.  Thus, even methods that seem to need only ensemble $N(z)$ may actually require accurate $p(z)$ in order to meet stringent survey requirements.
%In the case of the multiple types of probes of cosmology enabled by the
%LSST cosmology sample of several billion galaxies, the number of redshift bins and their photo-$z$ requirements vary with the specific probe;  2-point angular correlations benefit from many bins, while weak lensing probes do not (due to the wide lensing kernel).
Large photometric surveys such as LSST must develop algorithms that simultaneously meet the needs of all science cases.
In order to meet these ambitious goals for photo-$z$ accuracy, every aspect of photo-$z$ estimation will have to be optimized: the algorithms employed, both template and machine-learning based (both in design and implementation); the spectroscopic data used as a training set for machine learning algorithms or to estimate template sets and train Bayesian priors; and probabilistic catalogue compression schemes that balance information retention against limited storage resources.

There are numerous techniques for deriving photo-$z$ PDFs from photometry, yet no one method has yet been established as clearly superior.
Quantitative comparisons of photo-$z$ methods have been made before.
The Photo-$z$ Accuracy And Testing \citep[PHAT,][]{Hildebrandt:10} effort focused on point estimates derived from many photometric bands.  \citet{Rau:2015} introduced a new method for improving redshift PDFs using an ordinal classification algorithm.
DES compared several codes for point estimates and a subset with $p(z)$ information \citep{Sanchez:14}.  A follup up paper examined summary statistics of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}.  
This paper is distinguished by its focus on metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product.
The results presented in this work are
%part of the Data Challenge 1 (DC1)
a major focus of the Photometric Redshift working group of the LSST-DESC.
This work is laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}} as one of the critical activities to be completed in preparation for dark energy science analysis on the first year LSST data.
%This is the first of multiple papers by the working group, which will grow in sophistication.
In this initial paper we focus on evaluating the performance of photometric redshift codes and PDF-based performance metrics in the presence of representative training sets.
Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.  We note that these initial tests are a necessary requirement for photo-$z$ codes that will be used in cosmological analyses; however, meeting these requirements is only the first stage in the process, and can be thought of as an initial test under near perfect conditions before further complexities are added in future analyses.

Before moving forward, we must address how the best methods may be unique to the performance metrics and science cases considered and what distinguishes photo-$z$ PDFs of different methods from one another.
Though photo-$z$ PDFs are often written simply as $p(z)$, the PDF itself must be an interim posterior distribution $p(z | d, I)$, the probability of redshift conditioned on photometric data $d$ that has actually been observed and the prior information $I$ that guides how a redshift is extracted from the photometry.
If we run multiple photo-$z$ codes on a single dataset, the photo-$z$  posteriors will not be identical because each code employs assumptions that constitute an interim prior; these assumptions form the premise for photo-$z$ estimation as a whole and are the only way to introduce differences in estimates of what would otherwise be a shared photo-$z$ posterior $p(z | d)$ regardless of the code used to obtain it.
Though explicit knowledge of the interim prior is necessary to use photo-$z$ posteriors self-consistently in physical inference, the interim prior of a particular methodology is often implicit and not necessarily shared among all galaxies in the catalogue.

This paper therefore aims (1) to constrain the impact of the interim prior $I$ by separating it into a component $I_{H}$ representing the method itself and a component $I_{D}$ representing physical information, such as a training set or SED template library and (2) to present a procedure for evaluating the performance of photo-$z$ codes in generic tests that may include many more systematics in the interim prior $I$.
In order to isolate the effects encapsulated by $I_{H}$ of variation between codes from issues with the training set or template library encapsulated by $I_{D}$, we use an identical set of simulated galaxies for every code and construct a template library and training sample that are {\it complete and representative} and shared among all codes; that is, our training sample for machine learning codes is drawn from the same underlying galaxy population as our test set, with no additional selections, and the SED library used for template-based codes is the same as the one used to generate the photometric data.
We explore a number of performance metrics in this paper, in the controlled environment of the representative training data in order to establish a method for comparing photo-$z$ PDFs derived by different methods.
%not to make a conclusion regarding the superiority or even relative favorability of each code but to establish a method for comparing photo-$z$ PDFs derived by different methods.
These test conditions set the stage for addressing in a future paper the crucial issue of incomplete and non-representative prior information.
